### 2020-02-27 Writing/disparity_review.tex
Edits to the end

\noindent Morphological disparity analyses are widely employed in evolutionary palaeobiology, and they are based on a diversity of methods and data.
There is no ``one-size-fits-all'' pipeline for morphological disparity analyses.
As with any multidimensional analysis, there are many variables that have to be considered when deciding which data to use and how to analyse it, stemming from the explicit hypotheses being tested.
Many of the problems in morphological disparity analysis arise from ``blind'' application of established methodological pipelines without consideration of the biological question being addressed.
We advocate that researchers should assemble their analytical approaches based on an experimental approach that explores the impact of competing methods, such as choice of distance measure, ordination method and ancestral state estimation method on disparity analysis results.
Thankfully this is now getting easier through the availability of diverse well documented R packages for multidimensional analysis \citep{Navarro2003-vz, Bouxin2005-wk, oksanen2007vegan, Harmon2008-gq, lloyd2016, Guillerme2018-uc}.
Many of the methods employed in disparity analysis are used more widely in other fields, including genomics and ecology, which also encompass analyses of multidimensional datasets \citep{Donohue2013-bg, Saupe2015-vm, Canter2018-hk, mammola2019}.
Innovations in morphological disparity analyses likely await discovery in their respective literatures.

@@

Although disparity analyses are now simple to implement in freely available softwares \citep{Navarro2003-vz, Bouxin2005-wk, oksanen2007vegan, Harmon2008-gq, lloyd2016, Guillerme2018-uc} it is crucial to remember that they are multidimensional analyses and multidimensional analyses are complex.
We assert that future morphological analyses will benefit by emphasising the methodological decisions made, rather than simply using disparity analysis because \textit{we can}.

@@

This article results from discussion at the Royal Society International Science Seminar on `Reconciling disparate views on disparity' held at Chicheley Hall, 9-10th January, 2018.
AG was funded by European Research Council Starting grant 637171 ADaPTiVE;
ALJ was funded by an Irish Research Council Laureate Award IRCLA/2017/186;
ES was funded by a University of Adelaide Research Fellowship;
EES was funded by a Leverhulme Trust Research Project Grant (DGR01020);
PD was funded by NERC (NE/P013678/1; NE/N002067/1) and BBSRC (BB/N000919/1);
SB was funded by European Research Council (ERC) under the European Union's Horizon 2020 research and innovation program (grant agreement No 756226, ERC Starting Grant: PalM) and a Leverhulme Trust Research Project Grant (RPG-2017-167);
TG was funded by ARC DP170103227 and FT180100634 awarded to VW;

>>>>

\noindent Morphological disparity analyses are widely employed in evolutionary palaeobiology, and are based on a diversity of methods and data.
There is no ``one-size-fits-all'' pipeline for morphological disparity analyses.
As with any multidimensional analysis, there are many variables that have to be considered when deciding which data to use and how to analyse them, stemming from the explicit hypotheses being tested.
Many of the problems in morphological disparity analysis arise from ``blind'' application of established methodological pipelines without consideration of the biological question being addressed.
We advocate that researchers should assemble their analytical approaches based on an experimental approach that explores the impact of competing methods, such as choice of indices, ordination method and ancestral state estimation method on disparity analysis results.
Thankfully this is becoming easier through the availability of diverse well documented R packages for multidimensional analysis \citep{Navarro2003-vz, Bouxin2005-wk, oksanen2007vegan, Harmon2008-gq, lloyd2016, Guillerme2018-uc}.
Many of the methods employed in disparity analysis are used more widely in other fields, including genomics and ecology, which also encompass analyses of multidimensional datasets \citep{Donohue2013-bg, Saupe2015-vm, Canter2018-hk, mammola2019}.
Innovations in morphological disparity analyses likely await discovery in their respective literatures.

@@

Although disparity analyses are now simple to implement in freely available software \citep{Navarro2003-vz, Bouxin2005-wk, oksanen2007vegan, Harmon2008-gq, lloyd2016, Guillerme2018-uc} it is crucial to remember that they are multidimensional analyses and that multidimensional analyses are complex.
We assert that future morphological analyses will benefit from emphasising the methodological decisions made, rather than simply using disparity analysis because it exists.

@@

This article results from discussion at the Royal Society International Science Seminar on `Reconciling disparate views on disparity' held at Chicheley Hall, 9-10th January, 2018.
AG was funded by European Research Council Starting grant 637171 ADaPTiVE;
ALJ was funded by an Irish Research Council Laureate Award IRCLA/2017/186;
ES was funded by a University of Adelaide Research Fellowship;
EES was funded by a Leverhulme Trust Research Project Grant (DGR01020);
PD was funded by NERC (NE/P013678/1; NE/N002067/1) and BBSRC (BB/N000919/1);
SB was funded by European Research Council (ERC) under the European Union's Horizon 2020 research and innovation program (grant agreement No 756226, ERC Starting Grant: PalM) and a Leverhulme Trust Research Project Grant (RPG-2017-167);
TG was funded by ARC DP170103227 and FT180100634 awarded to VW.
